Carbonate chemistry at Ny-Ålesund

### 2021-02-04 awipev-CO2.Rmd
reorganized variable names and order of colmuns

```{r read observatory data, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
d_all <- as_tibble(d_hour) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth = ifelse( # create new variable depth
      pressure_insitu_ctd <= 2,
      1,
      ifelse(
        pressure_insitu_ctd > 2 & pressure_insitu_ctd <= 4,
        3,
        ifelse(
          pressure_insitu_ctd > 4 & pressure_insitu_ctd <= 6,
          5,
          ifelse(
            pressure_insitu_ctd > 6 & pressure_insitu_ctd <= 8,
            7,
            ifelse(pressure_insitu_ctd > 8, 9, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
```

@@

# Duplicate datetime in closest_datetime to do the left_join
d_all$closest_datetime <- d_all$datetime
discrete <- left_join(discrete, d_all %>% dplyr::select(c(closest_datetime, voltEXT, voltINT)), by="closest_datetime")
```

@@

```{r pco2 to fco2, echo=FALSE, warning=FALSE, message=FALSE}
d_all <- d_all %>%
  dplyr::mutate(
    fco2_fb = seacarb::p2fCO2(
      pCO2 = pco2_fb,
      T = temp_fb,
      Patm = 1,
      P = 0),
    fco2_fb_insi_temp = seacarb::fCO2insi(
      fCO2lab = fco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    .after = pco2_fb
)

@@

```{r save all, echo=FALSE, message= FALSE, warning=FALSE}
# d_hour is the real data set that is read at the beginning of the script
# d_all comprises additional columns calculated in this script
d_all$at <- ifelse(is.nan(d_all$at), NA, d_all$at) # replace all NaN by NA
d_all$ct <- ifelse(is.nan(d_all$at), NA, d_all$ct) # replace all NaN by NA
d_all <- dplyr::mutate(d_all, month = as.factor(month(x = datetime, label = TRUE)))
saveRDS(file= paste0(path, "fb_data/d_all.rds"), d_all)

>>>>

```{r read observatory data, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
d_all <- as_tibble(d_hour) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth = ifelse( # create new variable depth
      pressure_insitu_ctd <= 2,
      1,
      ifelse(
        pressure_insitu_ctd > 2 & pressure_insitu_ctd <= 4,
        3,
        ifelse(
          pressure_insitu_ctd > 4 & pressure_insitu_ctd <= 6,
          5,
          ifelse(
            pressure_insitu_ctd > 6 & pressure_insitu_ctd <= 8,
            7,
            ifelse(pressure_insitu_ctd > 8, 9, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
```

@@

# Duplicate datetime in closest_datetime to do the left_join
d_all$closest_datetime <- d_all$datetime
discrete <- left_join(discrete, d_all %>% dplyr::select(c(closest_datetime, voltEXT, voltINT)), by="closest_datetime")

# order columns
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all))) %>% 
  dplyr::select(datetime, date, year, monthd, montht, everything())
```

@@

```{r pco2 to fco2, echo=FALSE, warning=FALSE, message=FALSE}
d_all <- d_all %>%
  dplyr::mutate(
    fco2_fb = seacarb::p2fCO2(
      pCO2 = pco2_fb,
      T = temp_fb,
      Patm = 1,
      P = 0),
    fco2_fb_insi_temp = seacarb::fCO2insi(
      fCO2lab = fco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m)
)

@@

```{r save all, echo=FALSE, message= FALSE, warning=FALSE}
# d_hour is the real data set that is read at the beginning of the script
# d_all comprises additional columns calculated in this script
d_all$at <- ifelse(is.nan(d_all$at), NA, d_all$at) # replace all NaN by NA
d_all$ct <- ifelse(is.nan(d_all$at), NA, d_all$ct) # replace all NaN by NA
#d_all <- dplyr::mutate(d_all, month = as.factor(month(x = datetime, label = TRUE)))

# order columns
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all))) %>% 
  dplyr::select(datetime, date, year, monthd, montht, everything())
